### 2025-04-25 README.Rmd
change dev.period to dev_period; closes #50

# hatchR: Predict Fish Hatch and Emergence Timing <a href="https://bmait101.github.io/hatchR/"><img src="man/figures/logo.png" align="right" height="139" alt="hatchR website" /></a>

<!-- badges: start -->
[![R-CMD-check](https://github.com/bmait101/hatchR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bmait101/hatchR/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/hatchR)](https://CRAN.R-project.org/package=hatchR)
[![cran checks](https://badges.cranchecks.info/worst/hatchR.svg)](https://cran.r-project.org/web/checks/check_results_hatchR.html)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
<!-- badges: end -->

**hatchR** is an R package for predicting hatch and emergence timing for a wide range of wild fishes using the effective value framework (Sparks et al., (2019) <DOI:10.1139/cjfas-2017-0468>). 'hatchR' offers users access to established phenological models and the flexibility to incorporate custom parameterizations using external datasets. The package includes functions for model selection, phenology prediction, and visualization. **hatchR** is designed to be user-friendly and accessible to a wide range of users, including researchers, managers, and educators.

## Citation

If you use **hatchR** in a formal publication or report, please cite it. Citing **hatchR** lets us devote more resources to it in the future. View the **hatchR** citation by running:

```{r}
citation(package = "hatchR")
```

## Installation Instructions

Install the released version from CRAN:

@@

Install the development version of **hatchR** from GitHub:

@@

## Interactive Shiny App

@@

```{r}
# see days to hatch
WI_hatch$days_to_develop
# and development period
WI_hatch$dev.period
```

>>>>

# hatchR <a href="https://bmait101.github.io/hatchR/"><img src="man/figures/logo.png" align="right" height="139" alt="hatchR website" /></a>

[![publication](https://img.shields.io/badge/Cite-Unpublished-yellow)](https://github.com/bmait101/hatchR/blob/master/inst/CITATION)
[![downloads](https://cranlogs.r-pkg.org/badges/hatchR)](https://cran.r-project.org/package=hatchR)
[![total](https://cranlogs.r-pkg.org/badges/grand-total/hatchR)](https://cranlogs.r-pkg.org/)

***Predict hatch and emergence timing for wild fish***

---

**hatchR** is an R package for predicting hatch and emergence timing for a wide range of wild fishes using the effective value framework (Sparks et al., (2019) <DOI:10.1139/cjfas-2017-0468>). **hatchR** offers users access to established phenological models and the flexibility to incorporate custom parameterizations using external datasets. The package includes functions for model selection, phenology prediction, and visualization. **hatchR** is designed to be user-friendly and accessible to a wide range of users, including researchers, managers, and educators.

## Installation

[![CRAN](https://www.r-pkg.org/badges/version/hatchR)](https://CRAN.R-project.org/package=hatchR)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
<!-- [![R-CMD-check](https://github.com/bmait101/hatchR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bmait101/hatchR/actions/workflows/R-CMD-check.yaml) -->
<!-- [![cran checks](https://badges.cranchecks.info/worst/hatchR.svg)](https://cran.r-project.org/web/checks/check_results_hatchR.html) -->

Install the released version of **hatchR** from CRAN:

@@

Install the latest development version of **hatchR** from GitHub:

@@

## Interactive Shiny App <img src='https://upload.wikimedia.org/wikipedia/commons/b/bf/Shiny_hex_logo.svg' align="right" height="50" />

@@

```{r}
# see days to hatch
WI_hatch$days_to_develop
# and development period
WI_hatch$dev_period
```
